---
title: "SARS-2 viral kinetics"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup}
library(tidyverse)
library(deSolve)
library(mlxR)
```

# Load simulated treatments
```{r}
hcq.gautret.lim2009.sim.typical <- read.csv("data/hcq.gautret.lim2009.sim.typical.csv" )[-1]
hcq.who.lim2009.sim.typical <- read.csv("data/hcq.who.lim2009.sim.typical.csv")[-1]
covid.sim.ivmfood.3x600.typical <- read.csv("data/covid.sim.ivmfood.3x600.typical.csv")[-1]
covid.sim.ivmfood.3x300.typical <- read.csv("data/covid.sim.ivmfood.3x300.typical.csv")[-1]
lpvrtv.dickinson2011.sim.typical <- read.csv("data/lpvrtv.dickinson2011.sim.typical.csv")[-1]
ntz1200.balderas2011.sim.typical <- read.csv("data/ntz1200.balderas2011.sim.typical.csv")[-1]
ntz2900.balderas2011.sim.typical <- read.csv("data/ntz2900.balderas2011.sim.typical.csv")[-1]
art.birgersson.exp.typical <- read.csv("data/art.birgersson.exp.typical.csv")[-1]
```

# Setup
## Conversions
The viral kinetics models predict absolute viral loads. The authors reported the cycle threshold (Ct) value from real time polymerase chain reaction (rtPCR) assays. As these vary by laboratory and analytical conditions, we chose to correlate model output with observed Ct values with a regression model described by Chou et al., Clin Chem. 2020; 66: 549-555): Ct(load [log10]) = -3.502*load + 40.55

```{r}
Ct_viral_load <- function( Ct, viral_load ) {
  if( missing(Ct) )
    return (-3.502*viral_load+40.55)
  
  return( (Ct-40.55)/-3.502 )
}

Ct_viral_load( viral_load=3.0 )
Ct_viral_load( Ct=30 )
```

Positivity in rtPCR was set to a Ct value of 35.
```{r}
CT_LIMIT_OF_DETECTION <- 35
```

## Immunity
The response often follows a sigmoidal curve, similar to an Emax model. For example, anti-spike protein (RBD) IgG titres measured by [To et al., 2020](doi:10.1016/S1473-3099(20)30196-1) reach half-maximal levels after about 14 days.

### To et al., Lancet 2020
```{r echo=F}
IgG_RBD_To_EC50 <- 12
IgG_RBD_To_Emax <- 2.1
IgG_RBD_To_Hill <- 8
IgG_RBD_To_max_eps <- 42
IgG_RBD_response <- data.frame( days = 0:60,
                  IgG_RBD = ((0:60)^IgG_RBD_To_Hill)/(IgG_RBD_To_EC50^IgG_RBD_To_Hill+(0:60)^IgG_RBD_To_Hill))
```

### Long et al., Nat Med 2020
VS fit to seroconversion profiles (Long, Q., Liu, B., Deng, H. et al. Antibody responses to SARS-CoV-2 in patients with COVID-19. Nat Med 26, 845–848 (2020). https://doi.org/10.1038/s41591-020-0897-1)
```{r}
IgG_Long_EC50 <- 10.2 
IgG_Long_Emax <- 104.6 
IgG_Long_Hill <- 3.4
IgG_Long_max_eps <- 52
IgG_Long_response <- data.frame( days = 0:60,
                  IgG_Long = ((0:60)^IgG_Long_Hill)/(IgG_Long_EC50^IgG_Long_Hill+(0:60)^IgG_Long_Hill))
```

### Comparison of the two curves
```{r}
ggplot(IgG_Long_response %>% filter(days<40), aes(days, IgG_Long_Emax*IgG_Long) ) +
  geom_line( data=IgG_Long_response %>% filter(days<40), aes(days, IgG_Long_Emax*IgG_Long, color="Seroconv. (Long, 2020)" )) + 
  geom_line( data=IgG_RBD_response %>% filter(days<40), aes(days, IgG_Long_Emax*IgG_RBD, color="IgG RBD (To, 2020") ) +
  labs( title="Immune response",
        colour="Source",
        x="Time [d]",
        y="Converted [%]") +
  theme_minimal()
```

# PK models

## HCQ
Lim et al. (Antimicrobial Agents and Chemotherapy Mar 2009, 53 (4) 1468-1475; DOI: 10.1128/AAC.00339-08)

```{r echo=F}
hcq.lim2009.model <- inlineModel("
[INDIVIDUAL]
input = { Tlag_pop, ka_pop, Cl_pop, V1_pop, Q_pop, V2_pop,
          Tlag_sd, ka_sd, Cl_sd, V1_sd, Q_sd, V2_sd,
          prop, add,
          WT }

EQUATION:
covWT = log(WT) - 4.248495 ; centered on log(70)

DEFINITION:
Tlag  = { distribution=lognormal, typical=Tlag_pop,   sd=Tlag_sd }
Cl  =   { distribution=lognormal, typical=Cl_pop,     sd=Cl_sd }
V1  =   { distribution=lognormal, typical=V1_pop,     sd=V1_sd }
V2  =   { distribution=lognormal, typical=V2_pop,     sd=V2_sd }
Q  =    { distribution=lognormal, typical=Q_pop,      sd=Q_sd }
ka  =   { distribution=lognormal, typical=ka_pop,     sd=ka_sd }

[LONGITUDINAL]
input = {Tlag, ka, Cl, V1, Q, V2, add, prop}

EQUATION:
; Parameter transformations 
V = V1 
k = Cl/V1 
k12 = Q/V1 
k21 = Q/V2

; PK model definition
Cc = pkmodel(Tlag, ka, V, k, k12, k21)

OUTPUT:
output = {Cc, y}

DEFINITION:
y = {distribution=normal, prediction=Cc, errorModel=combined1(add, prop)}
")

HCQ_MW <- 433.95

HCQ_spike_EC50_Liu <- 8.51 
HCQ_unbound_fraction <- 0.5 # 50% of drug is bound to protein
```

## IVM
Duthaler et al., J Antimicrob Chemother. 2020 Feb 1;75(2):438-440.
doi: 10.1093/jac/dkz466

```{r echo=F}
ivmfood.model <- inlineModel("
[INDIVIDUAL]
input = { NN_pop, Mtt_pop, ka_pop, Cl_pop, V1_pop, Q_pop, V2_pop, F1_pop,
          NN_sd, Mtt_sd, ka_sd, Cl_sd, V1_sd, Q_sd, V2_sd,
          prop, add,
          WT }

EQUATION:
covWT = log(WT) - 4.248495 ; centered on log(70)
F1 = F1_pop

DEFINITION:
NN  = { distribution=lognormal, typical=NN_pop,   sd=NN_sd }
Mtt = { distribution=lognormal, typical=Mtt_pop,  sd=Mtt_sd }
ka  = { distribution=lognormal, typical=ka_pop,   sd=ka_sd }
Cl  = { distribution=lognormal, typical=Cl_pop,   sd=Cl_sd,   covariate=covWT,  coefficient=0.75 }
V1  = { distribution=lognormal, typical=V1_pop,   sd=V1_sd,   covariate=covWT,  coefficient=1 }
Q   = { distribution=lognormal, typical=Q_pop,    sd=Q_sd }
V2  = { distribution=lognormal, typical=V2_pop,   sd=V2_sd }

[LONGITUDINAL]
input = { NN, Mtt, ka, Cl, V1, Q, V2, F1, prop, add, WT}

EQUATION:
odeType = stiff
ddt_AUC = 1/V1 * Ac

; Parameter transformations 
Ktr = (NN+1)/Mtt
V = V1 
k = Cl/V1 
k12 = Q/V1 
k21 = Q/V2

; PK model definition
PK:
compartment(cmt=1, amount=Ac, volume=V)
peripheral( k12, k21 )
elimination( k )
oral( cmt=1, ka, Mtt, Ktr, p=F1 )
Cc = Ac / V

OUTPUT:
output = {Cc, y, AUC, WT}

DEFINITION:
y = {distribution=normal, prediction=Cc, errorModel=combined1(add, prop)}
")

IVM_MW <- 875.1 # g/mol, for conversion to uM

LUNG_ACCUMULATION <- 2.6 # cattle lung accumulation, Ref: Lifschitz
IVM_unbound_fraction <- (1-0.93) #93% of drug is protein bound

nAChR_IC50_uM <- 156 / 1000 #https://www.nature.com/articles/srep42481
nAChR_Hill <- 1 # Hill coefficient missing

# Helicase inhibition
# Ref: J Antimicrob Chemother. 2012 Aug;67(8):1884-94.
# doi: 10.1093/jac/dks147. Epub 2012 Apr 25. https://pubmed.ncbi.nlm.nih.gov/22535622/
YFV_IC50_uM <- 0.12 # +/- 0.01
DENV_IC50_uM <- 0.5 # +/- 0.07
WNV_IC50_uM <- 0.35 # +/- 0.04
HELICASE_IC50_uM <- 0.25
HELICASE_IC50_uM <- 0.1 # 10 fold higher IC50 in Vero, so 10 fold lower IC50 for helicase
```


## LPV/RTV
Dickinson L, Boffito M, Back D, et al. Sequential population pharmacokinetic modeling of lopinavir and ritonavir in healthy volunteers and assessment of different dosing strategies. Antimicrob Agents Chemother. 2011;55(6):2775-2782. doi:10.1128/AAC.00887-10

```{r echo=F}
lpvrtv.dickinson2011.model <- inlineModel("
[INDIVIDUAL]
input = {Tlag_pop, ka_pop, Cl_pop, V_pop,
         Tlag_sd, ka_sd, Cl_sd, V_sd,
         prop, add,
         WT}

DEFINITION:
Tlag = {distribution=lognormal, typical=Tlag_pop, sd=Tlag_sd}
ka =   {distribution=lognormal, typical=ka_pop,   sd=ka_sd}
Cl =   {distribution=lognormal, typical=Cl_pop,   sd=Cl_sd}
V =    {distribution=lognormal, typical=V_pop,    sd=V_sd}

[LONGITUDINAL]
input = {Tlag, ka, Cl, V, add, prop}

EQUATION:
; Parameter transformations 
k = Cl/V

; PK model definition
Cc = pkmodel(Tlag, ka, V, k)

OUTPUT:
output = {Cc}
")

LOP_MW <- 628.81 # g/mol, for conversion to uM

LUNG_ACCUMULATION <- 1.78
LOP_unbound_fraction <- (1-0.99)

Protease_IC50_uM <- 26.63 #uM Ref: Choy, K.-T., et al. (2020). "Remdesivir, lopinavir, emetine, and homoharringtonine inhibit SARS-CoV-2 replication in vitro." Antiviral research 178: 104786-104786.
Protease_Hill <- 1
```

## NTZ
```{r echo=F}
ntz.balderas2011.model <- inlineModel("
[INDIVIDUAL]
input = {Tlag_pop, ka_pop, Cl_pop, V_pop,
         Tlag_sd, ka_sd, Cl_sd, V_sd,
         WT}

DEFINITION:
Tlag = {distribution=lognormal, typical=Tlag_pop, sd=Tlag_sd}
ka =   {distribution=lognormal, typical=ka_pop,   sd=ka_sd}
Cl =   {distribution=lognormal, typical=Cl_pop,   sd=Cl_sd}
V =    {distribution=lognormal, typical=V_pop,    sd=V_sd}


[LONGITUDINAL]
input = {Tlag, ka, Cl, V}

EQUATION:
; Parameter transformations 
k = Cl/V

; PK model definition
Cc = pkmodel(Tlag, ka, V, k)

OUTPUT:
output = {Cc}
")

TZ_MW <- 265.25 # g/mol, for conversion to uM

LUNG_ACCUMULATION <- 0.7

TZ_unbound_fraction <- (1-0.99) 

Repl_IC50_uM <- 2.12 #uM Ref: Wang 2020
Protease_Hill <- 1
```


## ART
Malar J. 2016 Feb 16;15:90. doi: 10.1186/s12936-016-1134-8.
Population pharmacokinetic properties of artemisinin in healthy male Vietnamese volunteers
Sofia Birgersson, Pham Van Toi, Nguyen Thanh Truong, Nguyen Thi Dung, Michael Ashton, Tran Tinh Hien, Angela Abelö, Joel Tarning
PMID: 26879816 PMCID: PMC4754918 DOI: 10.1186/s12936-016-1134-8 
    
```{r}
art.birgersson.model <- inlineModel( "
[LONGITUDINAL]
input = {Ktr, Mtt, V, Cl}

EQUATION:
odeType = stiff
ka = Ktr

; PK model definition
Cc = pkmodel(Ktr, Mtt, ka, V, Cl)

OUTPUT:
output = Cc")

ART_MW <- 282.336 # g/mol

ART_EC50_uM_Nair <- 70 # uM Nair

ART_unbound_fraction <- (1-0.88) #88% of drug is protein bound, Sidhu et al., 1997
```

# Viral kinetics
## Model definitions
```{r}
TIME_POSITIVITY <- 5.4*24 # days post infection
TIME_PRESENTATION <- 10.2*24 # days post infection

get_ode_states_parameters <- function( immune_response="RBD_To" ) { 
  c <- 5.07 # virus clearance
  delta <- 0.54 # cell death
  p <- 10.18 # viral production rate
  I0 <- 0 # number of infected cells
  T0 <- 1*10**5 # number of target cells
  R0 <- 3.79 # within-host reproduction number
  V0 <- 1 # initial load (inoculum)
  
  Immune_EC50 <- IgG_RBD_To_EC50 # days
  Immune_Hill <- IgG_RBD_To_Hill # Hill coefficient
  Immune_Emax <- IgG_RBD_To_max_eps # maximum effect on viral clearance
  
  if( immune_response == "Sero_Long" ) {
    Immune_EC50 <- IgG_Long_EC50 # days
    Immune_Hill <- IgG_Long_Hill # Hill coefficient
    Immune_Emax <- IgG_Long_max_eps # maximum effect on viral clearance    
  }
  
  beta <- c *delta*R0/((p-delta*R0)*T0)

  # ODE system parameterization
  parameters <- c( c=c,
                   delta=delta,
                   p=p,
                   R0=R0,
                   beta=beta,
                   T0 = T0,
                   Immune_EC50=Immune_EC50,
                   Immune_Emax=Immune_Emax,
                   Immune_Hill=Immune_Hill)
  
  # ODE system initial conditions
  state <- c( Tf = T0, # target cells
              If = I0, # infected cells
              Vf = V0 ) # viral load
  
  return( list(state=state, parameters=parameters))
} 

get_ode_states_parameters()
get_ode_states_parameters("RBD_To")
get_ode_states_parameters("Sero_Long")
```

```{r echo=F}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# model
sars2_noeclipse <- function( t, state, parameters ) {  # function of time, state, parameters
  with(as.list(c(state, parameters)),{ 
    # rate of change
    eps_inf <- input_inf(t)
    eps_repro <- input_repro(t)
    eps_acquired <- Immune_Emax * (t^Immune_Hill)/(Immune_EC50^Immune_Hill+t^Immune_Hill)
    beta <- c*(1+eps_acquired)*delta*R0/((p-delta*R0)*T0)
    
    dTf <-  - beta*(1-eps_inf)*Tf*Vf
    dIf <-  + beta*(1-eps_inf)*Tf*Vf - delta*If
    dVf <-  + p*(1-eps_repro)*If - c*(1+eps_acquired)*Vf - beta*(1-eps_inf)*Tf*Vf
    # return rate of change
    list(c(dTf, dIf, dVf))
    })
}

# Immunity model according to Long
ode_settings <- get_ode_states_parameters("Sero_Long")

df.immunity.long <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="immune response (Long 2020)")

ggplot( df.immunity.long %>% filter(time<30), aes(time, Ct)) +
  geom_hline( yintercept = CT_LIMIT_OF_DETECTION, linetype="dotted", color="black", size=1 ) +
  geom_line(size=1) +
  scale_y_continuous( breaks=(0:10)*5, trans="reverse") +
  labs( color="",
        linetype="",
        x="Days post infection (dpi) [d]",
        y="Ct") + 
  theme_minimal() +
  theme( legend.position = "bottom")
```

```{r}
df.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="immune response (To 2020)")
```

## HCQ
### Gautret et al.
```{r}
hcq.lim2009.params <- data.frame(
  Tlag_pop = 0.389, Tlag_sd = 0,
  ka_pop = 1.15, ka_sd = 0,
  Cl_pop = 15.5, Cl_sd = 0,
  V1_pop = 733, V1_sd = 0,
  Q_pop = 45.1, Q_sd = 0,
  V2_pop = 1630, V2_sd=0,
  prop = 0, add = 0,
  WT = 70)

set.seed(1234)
study_h = 5*28*24
Cc <- list(name='Cc', time=seq(0, study_h, by=0.1))
adm <- list(amount = c(rep(400,5*4))*1000, time=seq(0, 3192, by=168))

# uncomment to rerun
# hcq.gautret.lim2009.sim.typical <- simulx(model = hcq.lim2009.model,
#                                   output    = list(Cc),
#                                   parameter = hcq.lim2009.params,
#                                   treatment = adm,
#                                   settings  = list(seed = 12342))

hcq.gautret.lim2009.sim.typical %<>% 
  mutate( CC_umol = Cc / HCQ_MW ) %>%   
  mutate( eps_spike_liu = CC_umol/(HCQ_spike_EC50_Liu+CC_umol),
          eps_unbound_liu = CC_umol*HCQ_unbound_fraction/(HCQ_spike_EC50_Liu+CC_umol*HCQ_unbound_fraction))

hcq.who.lim2009.sim.typical %<>% 
  mutate( CC_umol = Cc / HCQ_MW ) %>%   
  mutate( eps_spike_liu = CC_umol/(HCQ_spike_EC50_Liu+CC_umol),
          eps_unbound_liu = CC_umol*HCQ_unbound_fraction/(HCQ_spike_EC50_Liu+CC_umol*HCQ_unbound_fraction))
```

#### Dosing on positivity
```{r echo=F}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# Gautret treatment effect
df <- hcq.gautret.lim2009.sim.typical %>% 
  dplyr::mutate( time = time + TIME_POSITIVITY ) 
df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_spike_liu=0,
                         eps_unbound_liu=0 ),
             df ) 
input_inf <- approxfun( df$time/24, df$eps_unbound_liu) # effect of infectivity suppression

df.gautret.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="HCQ Gautret, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.gautret.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, HCQ 200 mg q8h (10d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r echo=F}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# Gautret treatment effect
df <- hcq.gautret.lim2009.sim.typical %>% 
  mutate( time = time + TIME_PRESENTATION )
df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_spike_liu=0,
                         eps_unbound_liu=0),
             df ) 
input_inf <- approxfun( df$time/24, df$eps_unbound_liu) # effect of infectivity suppression

df.gautret.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="HCQ Gautret, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.gautret.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line(size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, HCQ 200 mg q8h (10d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

### WHO Solidarity
```{r}
hcq.lim2009.params <- data.frame(
  Tlag_pop = 0.389, Tlag_sd = 0,
  ka_pop = 1.15, ka_sd = 0,
  Cl_pop = 15.5, Cl_sd = 0,
  V1_pop = 733, V1_sd = 0,
  Q_pop = 45.1, Q_sd = 0,
  V2_pop = 1630, V2_sd=0,
  prop = 0.27, add = 2.77,
  WT = 70)

set.seed(1234)
study_h = 40*24
y <- list(name='y', time=seq(0, study_h, by=0.1))
Cc <- list(name='Cc', time=seq(0, study_h, by=0.1))
adm <- list(amount = c(800, 800, rep(400,9*2))*1000, time=c(0:19)*12)

# uncomment to rerun
# hcq.who.lim2009.sim.typical <- simulx(model     = hcq.lim2009.model,
#                                   output    = list(Cc, y),
#                                   parameter = hcq.lim2009.params,
#                                   treatment = adm,
#                                   settings  = list(seed = 12342))

hcq.who.lim2009.sim.typical %<>% 
  mutate( CC_umol = Cc / HCQ_MW ) %>%   
  mutate( eps_spike_liu = CC_umol/(HCQ_spike_EC50_Liu+CC_umol),
          eps_unbound_liu = CC_umol*HCQ_unbound_fraction/(HCQ_spike_EC50_Liu+CC_umol*HCQ_unbound_fraction))

hcq.who.lim2009.sim.typical %<>% 
  mutate( CC_umol = Cc / HCQ_MW ) %>%   
  mutate( eps_spike_liu = CC_umol/(HCQ_spike_EC50_Liu+CC_umol),
          eps_unbound_liu = CC_umol*HCQ_unbound_fraction/(HCQ_spike_EC50_Liu+CC_umol*HCQ_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# Gautret treatment effect
df <- hcq.who.lim2009.sim.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_spike_liu=0,   
                         eps_unbound_liu=0 ),
             df ) 
input_inf <- approxfun( df$time/24, df$eps_unbound_liu) # effect of infectivity suppression

df.who.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="HCQ WHO, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.who.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, WHO)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# Gautret treatment effect
df <- hcq.who.lim2009.sim.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_spike_liu=0,                         
                         eps_unbound_liu=0),
             df ) 
input_inf <- approxfun( df$time/24, df$eps_unbound_liu) # effect of infectivity suppression

df.who.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="HCQ WHO, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.who.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, WHO)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```


## IVM
### 3x300
```{r}
covid.ivmfood.params <- data.frame(
  id = 1,
  Cl_pop = 7.74, Cl_sd = 0,
  V1_pop = 100.8, V1_sd = 0,
  V2_pop = 228.8, V2_sd = 0,
  Q_pop = 18.67, Q_sd = 0,
  ka_pop = 0.68, ka_sd = 0,
  Mtt_pop = 0.93, Mtt_sd = 0,
  NN_pop = 6, NN_sd = 0,
  F1_pop = 1, # fed
  prop = 0.09, 
  add = 0.82,
  WT = 70
)

adm  <- data.frame(id=1,
                   amount=rep(300*70, 3), #300ug/kg for a 70kg individual: 21mg
                   time = c(0, 24, 48))

# uncomment to rerun
# covid.sim.ivmfood.3x300.typical <- simulx( model = ivmfood.model,
#                                            output = list(Cc, y),
#                                            parameter = covid.ivmfood.params,
#                                            treatment = adm )

# 3x300 IVM helicase and nAChR effect
covid.sim.ivmfood.3x300.typical %<>% 
  mutate( CC_umol = Cc / IVM_MW ) %>%  
  mutate( eps_nAChR = CC_umol/(nAChR_IC50_uM+CC_umol),
          eps_nAChR_lung = LUNG_ACCUMULATION*CC_umol/(nAChR_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          # eps_nAChR_unbound = IVM_unbound_fraction*CC_umol/ (nAChR_IC50_uM+CC_umol*IVM_unbound_fraction),
          eps_nAChR_unbound = LUNG_ACCUMULATION*IVM_unbound_fraction*CC_umol/ (nAChR_IC50_uM+LUNG_ACCUMULATION*CC_umol*IVM_unbound_fraction),
          eps_helicase = CC_umol/(HELICASE_IC50_uM+CC_umol),
          eps_helicase_lung = LUNG_ACCUMULATION*CC_umol/(HELICASE_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          eps_helicase_unbound = CC_umol*IVM_unbound_fraction/ (HELICASE_IC50_uM+CC_umol*IVM_unbound_fraction))

                    # eps_helicase_unbound = LUNG_ACCUMULATION*CC_umol*IVM_unbound_fraction/ (HELICASE_IC50_uM+LUNG_ACCUMULATION*CC_umol*IVM_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# 3x300 treatment effect
df <- covid.sim.ivmfood.3x300.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_nAChR=0,
                         eps_nAChR_lung=0,
                         eps_nAChR_unbound=0,
                         eps_helicase=0,
                         eps_helicase_lung=0,
                         eps_helicase_unbound=0),
             df ) 
input_repro <- approxfun( df$time/24, df$eps_helicase_unbound) # effect of helicase suppression
input_inf <- approxfun( df$time/24, df$eps_nAChR_unbound) # effect of infectivity suppression

df.3x300.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="IVM 3x300, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.3x300.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, IVM 3x300 ug/kg",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# 3x300 treatment effect
df <- covid.sim.ivmfood.3x300.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d
df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_nAChR=0,
                         eps_nAChR_lung=0,
                         eps_nAChR_unbound=0,
                         eps_helicase=0,
                         eps_helicase_lung=0,
                         eps_helicase_unbound=0),
             df )
input_repro <- approxfun( df$time/24, df$eps_helicase_unbound) # effect of helicase suppression
input_inf <- approxfun( df$time/24, df$eps_nAChR_unbound) # effect of infectivity suppression

df.3x300.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="IVM 3x300, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.3x300.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, IVM 3x300 ug/kg",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

### 3x600
```{r}
covid.ivmfood.params <- data.frame(
  id = 1,
  Cl_pop = 7.74, Cl_sd = 0,
  V1_pop = 100.8, V1_sd = 0,
  V2_pop = 228.8, V2_sd = 0,
  Q_pop = 18.67, Q_sd = 0,
  ka_pop = 0.68, ka_sd = 0,
  Mtt_pop = 0.93, Mtt_sd = 0,
  NN_pop = 6, NN_sd = 0,
  F1_pop = 1, # fed
  prop = 0.09, 
  add = 0.82,
  WT = 70
)

adm  <- data.frame(id=1,
                   amount=rep(600*70, 3), #600ug/kg for a 70kg individual: 42mg
                   time = c(0, 24, 48))

# uncomment to rerun
# covid.sim.ivmfood.3x600.typical <- simulx( model = ivmfood.model,
#                                            output = list(Cc, y),
#                                            parameter = covid.ivmfood.params,
#                                            treatment = adm )

# 3x600 IVM helicase and nAChR effect
covid.sim.ivmfood.3x600.typical %<>% 
  mutate( CC_umol = Cc / IVM_MW ) %>% 
  mutate( eps_nAChR = CC_umol/(nAChR_IC50_uM+CC_umol),
          eps_nAChR_lung = LUNG_ACCUMULATION*CC_umol/(nAChR_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          # eps_nAChR_unbound = IVM_unbound_fraction*CC_umol/ (nAChR_IC50_uM+CC_umol*IVM_unbound_fraction),
          eps_nAChR_unbound = LUNG_ACCUMULATION*IVM_unbound_fraction*CC_umol/ (nAChR_IC50_uM+LUNG_ACCUMULATION*CC_umol*IVM_unbound_fraction),
          eps_helicase = CC_umol/(HELICASE_IC50_uM+CC_umol),
          eps_helicase_lung = LUNG_ACCUMULATION*CC_umol/(HELICASE_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          # eps_helicase_unbound = CC_umol*IVM_unbound_fraction/ (HELICASE_IC50_uM+CC_umol*IVM_unbound_fraction))
                    eps_helicase_unbound = LUNG_ACCUMULATION*CC_umol*IVM_unbound_fraction/ (HELICASE_IC50_uM+LUNG_ACCUMULATION*CC_umol*IVM_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# 3x600 treatment effect
df <- covid.sim.ivmfood.3x600.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_nAChR=0,
                         eps_nAChR_lung=0,
                         eps_nAChR_unbound=0,
                         eps_helicase=0,
                         eps_helicase_lung=0,
                         eps_helicase_unbound=0),
             df )
input_repro <- approxfun( df$time/24, df$eps_helicase_unbound) # effect of helicase suppression
input_inf <- approxfun( df$time/24, df$eps_nAChR_unbound) # effect of infectivity suppression

df.3x600.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="IVM 3x600, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.3x600.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, IVM 3x600 ug/kg",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_inf <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# 3x600 treatment effect
df <- covid.sim.ivmfood.3x600.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d


df <- rbind( data.frame( time=0, #add rows with zeros
                         Cc=0,
                         CC_umol=0,
                         eps_nAChR=0,
                         eps_nAChR_lung=0,
                         eps_nAChR_unbound=0,
                         eps_helicase=0,
                         eps_helicase_lung=0,
                         eps_helicase_unbound=0),
             df )
input_repro <- approxfun( df$time/24, df$eps_helicase_unbound) # effect of helicase suppression
input_inf <- approxfun( df$time/24, df$eps_nAChR_unbound) # effect of infectivity suppression

df.3x600.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="IVM 3x600, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.3x600.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, IVM 3x600 ug/kg",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

## LPV/RTV
```{r}
lpvrtv.dickinson2011.params <- data.frame(
  Tlag_pop = 0.7, Tlag_sd = 0,
  ka_pop = 0.26, ka_sd = 0,
  Cl_pop = 4.5, Cl_sd = 0,
  V_pop = 15.9, V_sd = 0,
  prop = 0.40, add = 0,
  WT = 70)

set.seed(1234)
study_h = 40*24
y <- list(name='y', time=seq(0, study_h, by=0.1))
Cc <- list(name='Cc', time=seq(0, study_h, by=0.1))
adm <- list(amount = c(rep(400))*1000, # mg converted in ug
            time=c(0:29)*12) # twice daily for 14 days

# uncomment to rerun
# lpvrtv.dickinson2011.sim.typical <- simulx(model = lpvrtv.dickinson2011.model,
#                                   output    = list(Cc),
#                                   parameter = lpvrtv.dickinson2011.params,
#                                   treatment = adm,
#                                   settings  = list(seed = 1234))

# protease effect
lpvrtv.dickinson2011.sim.typical %<>% 
  mutate( CC_umol = Cc / LOP_MW ) %>% 
  mutate( eps_protease = CC_umol/(Protease_IC50_uM+CC_umol), #Emax model
          eps_protease_lung = LUNG_ACCUMULATION*CC_umol/(Protease_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          eps_protease_unbound = LUNG_ACCUMULATION*CC_umol* LOP_unbound_fraction/ (Protease_IC50_uM+LUNG_ACCUMULATION*CC_umol*LOP_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# lpvrtv treatment effect
df <- lpvrtv.dickinson2011.sim.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_protease=0,
                         eps_protease_lung=0,
                         eps_protease_unbound=0),
             df ) 


input_inf <- approxfun( df$time/24, df$eps_protease_unbound) # effect of infectivity suppression

df.lpvrtv.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="LPV/r, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.lpvrtv.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, Lopinavir/Ritonavir q12h (14d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# lpvrtv treatment effect
df <- lpvrtv.dickinson2011.sim.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_protease=0,
                         eps_protease_lung=0,
                         eps_protease_unbound=0),
             df ) 
input_inf <- approxfun( df$time/24, df$eps_protease_unbound) # effect of infectivity suppression

df.lpvrtv.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="LPV/r, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.lpvrtv.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, Lopinavir/Ritonavir q12h (14d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```


## NTZ
### 1200 mg
```{r}
ntz1200.balderas2011.params <- data.frame(
  Tlag_pop = 0.271, Tlag_sd = 0,
  ka_pop = 0.415, ka_sd = 0,
  Cl_pop = 11.19, Cl_sd = 0,
  V_pop = 12.57, V_sd = 0,
  WT = 70)

set.seed(1234)
study_h = 40*24 
y <- list(name='y', time=seq(0, study_h, by=0.1))
Cc <- list(name='Cc', time=seq(0, study_h, by=0.1))
adm <- list(amount = c(rep(1200)), # ug
            time=c(0:19)*6) # 4 times daily for 5 days

# uncomment to rerun
# ntz1200.balderas2011.sim.typical <- simulx(model = ntz1200.balderas2011.model,
#                                   output    = list(Cc),
#                                   parameter = ntz1200.balderas2011.params,
#                                   treatment = adm,
#                                   settings  = list(seed = 1234))

ntz1200.balderas2011.sim.typical %<>% 
  mutate( CC_umol = Cc / TZ_MW ) %>% 
  mutate( eps_repl = CC_umol/(Repl_IC50_uM+CC_umol), #Emax model
          eps_repl_lung = LUNG_ACCUMULATION*CC_umol/(Repl_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          eps_repl_unbound = LUNG_ACCUMULATION*CC_umol* TZ_unbound_fraction/ (Repl_IC50_uM+LUNG_ACCUMULATION*CC_umol*TZ_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# NTZ treatment effect
df <- ntz1200.balderas2011.sim.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_repl=0,
                         eps_repl_lung=0,
                         eps_repl_unbound=0),
             df ) 


input_inf <- approxfun( df$time/24, df$eps_repl_unbound ) # effect of infectivity suppression

df.ntz1200.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="NTZ 1200mg, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.ntz1200.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, Nitazoxanide 1200 mg q6h (5d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# NTZ treatment effect
df <- ntz1200.balderas2011.sim.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_repl=0,
                         eps_repl_lung=0,
                         eps_repl_unbound=0),
             df ) 

input_repro <- approxfun( df$time/24, df$eps_repl_unbound ) # effect of infectivity suppression
input_inf <- approxfun( df$time/24, df$eps_repl_unbound ) # effect of infectivity suppression

df.ntz1200.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="NTZ 1200mg, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.ntz1200.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, Nitazoxanide 1200 mg q6h (5d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

### 2900 mg
```{r}
ntz2900.balderas2011.params <- data.frame(
  Tlag_pop = 0.271, Tlag_sd = 0,
  ka_pop = 0.415, ka_sd = 0,
  Cl_pop = 11.19, Cl_sd = 0,
  V_pop = 12.57, V_sd = 0,
  WT = 70)

set.seed(1234)
study_h =40*24
y <- list(name='y', time=seq(0, study_h, by=0.1))
Cc <- list(name='Cc', time=seq(0, study_h, by=0.1))
adm <- list(amount = c(rep(2900)), # ug
            time=c(0:9)*12) # 2 times daily for 5 days

# uncomment to rerun
# ntz2900.balderas2011.sim.typical <- simulx(model = ntz2900.balderas2011.model,
#                                   output    = list(Cc),
#                                   parameter = ntz2900.balderas2011.params,
#                                   treatment = adm,
#                                   settings  = list(seed = 1234))

ntz2900.balderas2011.sim.typical %<>% 
  mutate( CC_umol = Cc / TZ_MW ) %>% 
  mutate( eps_repl = CC_umol/(Repl_IC50_uM+CC_umol), #Emax model
          eps_repl_lung = LUNG_ACCUMULATION*CC_umol/(Repl_IC50_uM+LUNG_ACCUMULATION*CC_umol),
          eps_repl_unbound = LUNG_ACCUMULATION*CC_umol* TZ_unbound_fraction/ (Repl_IC50_uM+LUNG_ACCUMULATION*CC_umol*TZ_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# NTZ treatment effect
df <- ntz2900.balderas2011.sim.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_repl=0,
                         eps_repl_lung=0,
                         eps_repl_unbound=0),
             df ) 


input_inf <- approxfun( df$time/24, df$eps_repl_unbound) # effect of infectivity suppression

df.ntz2900.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="NTZ 2900mg, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.ntz2900.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, Nitazoxanide 2900 mg q12h (5d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# TZ treatment effect
df <- ntz2900.balderas2011.sim.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_repl=0,
                         eps_repl_lung=0,
                         eps_repl_unbound=0),
             df ) 
input_inf <- approxfun( df$time/24, df$eps_repl_unbound) # effect of infectivity suppression

df.ntz2900.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="NTZ 2900mg, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.ntz2900.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, Nitazoxanide 2900 mg q12h (5d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]") +
  theme_ipsum_ps()

ggplot(df.ntz2900.on_presentation.immunity %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, Nitazoxanide 2900 mg q12h (5d)",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```


## ART
```{r}
art.birgersson.params <- data.frame(
  Cl = 417, 
  V = 1210,
  Mtt = 0.787, 
  Ktr = (7+1)/0.787 # nb transit compartments = NN = 7 and Ktr = (NN+1)/Mtt
)

set.seed(1234)
study_h = 40*24
Cc <- list(name='Cc', time=seq(0, study_h, by=0.1))
adm <- list(amount = 500*1000, # 500 mg in ug
            time=c(0:4)*24) # every day for 5 days

# uncomment to rerun
# art.birgersson.exp.typical <- simulx( model = art.birgersson.model,
#                                            output = list(Cc),
#                                            parameter = art.birgersson.params,
#                                            treatment = adm )

art.birgersson.exp.typical %<>% 
  mutate(CC_umol = Cc/ART_MW) %>% 
  mutate(eps_art1 = CC_umol/(ART_EC50_uM_Nair+CC_umol), # 70 uM
         eps_art1_unbound = CC_umol*ART_unbound_fraction/(ART1_EC50_uM+CC_umol*ART_unbound_fraction))
```

#### Dosing on positivity
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# ART treatment effect
df <- art.birgersson.exp.typical %>% 
  mutate( time = time + TIME_POSITIVITY ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_art1_unbound=0,
                         eps_art1=0),
             df )

input_inf <- input_repro <- approxfun( df$time/24, df$eps_art1_unbound) 

df.art.on_positivity.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="ART 500 mg, treated on positivity")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.art.on_positivity.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on positivity, Artemisinin 500 mg qd 5d",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```

#### Dosing on presentation
```{r}
ode_settings <- get_ode_states_parameters(immune_response = "Sero_Long")

times <- seq(0, 40, # number of days
             by = .1) 

input_repro <- input_cl <- input_repro <- function(x) return(0) # set all effects to 0 at any time

# ART treatment effect
df <- art.birgersson.exp.typical %>% 
  mutate( time = time + TIME_PRESENTATION ) # shift right by 5d

df <- rbind( data.frame( time=0, 
                         Cc=0,
                         CC_umol=0,
                         eps_art1_unbound=0,
                         eps_art1=0),
             df )

input_inf <- input_repro <- approxfun( df$time/24, df$eps_art1_unbound) 

df.art.on_presentation.immunity <- ode(y = ode_settings$state, 
           times = times, 
           func = sars2_noeclipse, 
           parms = ode_settings$parameters) %>% 
  as.data.frame() %>% 
  mutate( Ct = Ct_viral_load( viral_load=log(Vf,base=10) ),
          group="ART 500 mg, treated on presentation")

ggplot( rbind(df.immunity %>% mutate(group="untreated"), 
              df.art.on_presentation.immunity) %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, Artemisinin 500 mg qd 5d",
        x="Days post infection [dpi]",
        y="Viral load [Ct]") 

ggplot(df.art.on_presentation.immunity %>% 
          filter( time<25), aes(time, Ct, group=group, color=group)) +
  geom_line( size=1) + 
  scale_y_reverse() +
  geom_hline( yintercept = 35, linetype="dashed", color="black") +
  labs( title= "Within-host dynamics of SARS-CoV-2",
        subtitle = "Dosing on presentation, Artemisinin 500 mg qd 5d",
        x="Days post infection [dpi]",
        y="Viral load [Ct]")
```